---
  title: "Exploring the BRFSS data"
output: 
  html_document: 
  fig_height: 4
highlight: pygments
theme: spacelab
---
  
  ## Setup
  
  ### Load packages
  
  ```{r load-packages, message = FALSE}
library(ggplot2)
library(dplyr)
```

### Load data

```{r load-data}
load("brfss2013.RData")
```



* * *
  
  ## Part 1: Data
  
  * * *
  
  ## Part 2: Research questions
  
  **Research question 1: Is there an observable difference in days where army veterans suffer poor health-related quality of life compared to non-veterans?**
  
  **Research question 2: Is there an observable difference in alcohol consumption observed in army veterans compared to the general population?**
  
  **Research question 3: Are there any significant differences in tobacco use in army veterans compared to the general population?**
  
  
  * * *
  
  ## Part 3: Exploratory data analysis
  
  NOTE: Insert code chunks as needed by clicking on the "Insert a new code chunk" 
button (green button with orange arrow) above. Make sure that your code is visible
in the project you submit. Delete this note when before you submit your work.

**Research question 1:**
  
  ```{r}
library(dplyr)
brfss2013 %>%
  filter(!is.na(menthlth),!is.na(physhlth),!is.na(poorhlth)) %>%
  group_by(veteran3) %>%
  summarise(poormentalhealthmean = mean(menthlth),poorphyshealthmean = mean(physhlth),poorhealthmean = mean(poorhlth))
```
Veterans seem to experience, on average, more days with poor quality of physical health than non-veterans; Mean days of poor mental health quality seemed to be indifferent between the two groups.


**Research question 2:**
  
  ```{r}
brfss2013 %>%
  filter(!is.na(avedrnk2),!is.na(maxdrnks),!is.na(alcday5)) %>%
  group_by(veteran3) %>%
  summarise(avgdrinks = mean(avedrnk2),meanmostdrinks = mean(maxdrnks),drinkdays = mean(alcday5))
```
There seems to be no evidence of veterans drinking more than non-veterans - if anything, veterans seem to have less drink days on average than non-veterans. 


**Research question 3:**
  
  ```{r}
brfss2013 %>%
  group_by(veteran3) %>%
  summarise(count=n())
brfss2013 %>%
  filter(smoke100 == "Yes", smokday2 == "Every day") %>%
  group_by(veteran3) %>%
  summarise(count=n())
```
The ratio of veterans:non-veterans seems to be roughly equal even after filtering for people who have smoked at least 100 cigarettes and who smoke everyday. 